% ============================================================================
% SUPERSEDED DRAFT. The current, canonical paper is paper/oicc_paper.tex (OICC).
% This older feedback-law draft contains a claim now known to be FALSE: that a
% difference-in-differences POINT-IDENTIFIES the loop gain kappa. It does not --
% the DiD carries an un-cancelled unobserved-latent-level term; only the
% recording response rho is identified (under parallel trends), and the real-data
% DiD is a null. Do NOT submit this draft. See docs/PROOFS_feedback_law.md
% (correction) and paper/oicc_paper.tex (the honest submission). Kept for record.
% ============================================================================
% Honest Forecasting under Observation-Biased Feedback
% Feedback-corrected conformal prediction for the latent process
%
% Target venues: FAccT 2027 / NeurIPS 2026 / a statistics-ML venue
%
% NOTE (positioning): the ZINB/GNN/conformal machinery is APPLIED prior art
% (STZINB-GNN Zhuang KDD'22; STMGNN-ZINB Wang'24; ACI Gibbs-Candes'21; CQR
% Romano'19). The contribution is the feedback-corrected LATENT conformal
% predictor + its DiD identification. See docs/NOVELTY_AND_POSITIONING.md.

\documentclass[10pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{algorithm, algorithmic}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

\title{\textbf{Confidently Wrong: Feedback-Corrected Conformal Prediction
for the Latent Process under Observation-Biased Allocation}}

\author{Himanshu Bairwa \\ \texttt{himanshu@example.edu}}
\date{}

\begin{document}
\maketitle

% ============================================================
\begin{abstract}
When a forecaster's predictions drive where attention is allocated, and
attention determines what gets recorded, the recorded data ceases to reflect the
latent process the system claims to predict. We study this
\emph{allocation under observation-biased feedback} (AOBF) setting and prove a
sharp and counterintuitive failure mode: as the feedback gain rises, a
correctly calibrated predictor \emph{maintains} coverage of the recorded process
while its coverage of the \emph{true latent} process silently collapses --- it
becomes \emph{confidently wrong}, with every observable metric improving as the
system becomes less trustworthy.

We make this precise. At the feedback fixed point the recorded intensity relates
to the latent intensity by an amplification elasticity
$d\log\mu / d\log\lambda = 1/(1-\kappa)$, where the feedback gain $\kappa$ is the
product of the policy and detection log-elasticities; the map has a pole at
$\kappa=1$ (runaway). This closed form is the classical closed-loop gain
(the social multiplier of Glaeser et al., 2003); our contribution is its
coordinate-free specialization to an observation-biased Poisson recording loop
and a group-disparity corollary that gives Ensign et al.'s (2018) runaway
feedback an exact functional form. We further show the confidently-wrong state
is \emph{not identifiable from passive data} but $\kappa$ is
\emph{point-identified} by a difference-in-differences on log recorded rates
after an exogenous detection-sensitivity shock (a staggered ShotSpotter/patrol
rollout).

Finally --- the constructive step no prior work provides --- we give a
\emph{feedback-corrected conformal predictor}: deflate the record by the
identified $\kappa$ and issue intervals valid for the latent process, abstaining
where correction is untrustworthy. In closed-loop simulation, naive latent
coverage collapses from $0.95$ to $0.16$ as feedback rises while our corrected
intervals hold at $0.93$--$0.96$ throughout. Prior work \emph{diagnoses} this
pathology; we \emph{correct} it. All code, proofs, and experiments are released.
\end{abstract}

% ============================================================
\section{Introduction}

Crime forecasts, infrastructure-inspection schedules, fraud audits, and
health-screening programs share a structure: predictions drive where limited
attention goes, and attention determines what is subsequently observed. Records
are therefore not samples of the world but of the world \emph{as the system
chose to look at it}. Ensign et al.~\cite{ensign2018} showed such loops can run
away in predictive policing; van Amsterdam et al.~\cite{vanamsterdam2025} showed
accurate models can be self-fulfillingly harmful in ways passive validation
cannot detect. Both \emph{diagnose}. Neither \emph{corrects}.

\textbf{This paper.} We formalize allocation under observation-biased feedback
(AOBF), give a closed-form amplification law and its identification, and provide
the first \emph{correction}: a conformal predictor whose intervals are valid for
the latent process, not the biased record.

\textbf{Contributions.}
\begin{enumerate}
  \item \textbf{The confidently-wrong phenomenon (\S\ref{sec:law}).} A closed-form
        amplification elasticity $1/(1-\kappa)$ at the AOBF fixed point, with a
        pole at $\kappa=1$; and a proof that as $\kappa$ rises, observed coverage
        is maintained while latent coverage collapses. We position this
        explicitly as a coordinate-free specialization of the classical
        closed-loop gain~\cite{glaeser2003}, sharpening Ensign
        et al.~\cite{ensign2018}.
  \item \textbf{Passive/active identification (\S\ref{sec:ident}).} The
        confidently-wrong state is non-identifiable from passive data; $\kappa$
        is point-identified by a difference-in-differences on log recorded rates
        after an exogenous detection shock. (The duality principle is
        known~\cite{mendlerduenner2022}; our instantiation for a recording-loop
        elasticity is new.)
  \item \textbf{Feedback-corrected conformal prediction (\S\ref{sec:correct}).}
        The constructive contribution: deflate the record by the identified
        $\kappa$, issue latent-valid intervals, abstain near runaway. Verified in
        closed loop (\S\ref{sec:exp}): latent coverage restored from $0.16$ to
        $0.95$.
  \item \textbf{A deployable safety layer (\S\ref{sec:routing}, \S\ref{sec:deploy}).}
        Feedback-aware routing that plans on de-biased latent risk and shrinks
        exposure disparity; an \emph{anytime-valid} tripwire (test
        supermartingale) that fires when latent coverage drifts, with a
        time-uniform false-alarm guarantee; and a robustness envelope bounding
        how much $\kappa$-misspecification the correction tolerates.
  \item \textbf{An open AOBF benchmark} and a released, tested implementation.
\end{enumerate}

We are explicit about what is \emph{not} new (the $1/(1-\kappa)$ form, the loop
itself, the identification-duality principle) and what is (the coordinate-free
recording-loop specialization, the DiD identification of $\kappa$, and the
correction). See \S\ref{sec:related}.

% ============================================================
\section{Related Work and Honest Positioning}
\label{sec:related}

\textbf{The amplification form is classical.} The closed-loop gain $1/(1-\kappa)$
is the social multiplier~\cite{glaeser2003} and the textbook feedback-amplifier
formula. We claim only its coordinate-free specialization ($\kappa$ as a product
of a policy log-elasticity and a detection log-elasticity, invariant to smooth
reparametrization) and the group-disparity power law
$b^{1/(1-\kappa)}$.

\textbf{The policing feedback loop is Ensign et al.~\cite{ensign2018}.} They give
a P\'olya-urn model with winner-take-all runaway; we give a smooth power law with
a finite pole, an elasticity decomposition, identification, and correction ---
none of which appear there.

\textbf{The identification duality is a known template}
\cite{mendlerduenner2022}: feedback parameters need exogenous variation. Our
contribution is the specific DiD-on-recorded-rates instantiation for the
recording-loop gain, distinct from reduced-form ShotSpotter DiD studies that
estimate treatment effects, not $\kappa$.

\textbf{Conformal-under-feedback controls the observed outcome.} ACI
\cite{gibbs2021}, Conformal PID, and Performative Risk Control all maintain
coverage/risk of the \emph{observed} target; none addresses a latent target
distinct from the record, and none corrects. The base predictor (ZINB + GNN) is
prior art~\cite{zhuang2022,wang2024}; we use it, not claim it.

% ============================================================
\section{The Confidently-Wrong Phenomenon}
\label{sec:law}

\textbf{Model (AOBF).} Latent intensity $\lambda_s>0$ (unobserved). Attention
$a_s=\pi(\mu_s;M)$ from the recorded-rate estimate $\mu_s$, $M=\text{mean}(\mu)$,
$\pi$ $C^1$ increasing. Observation-biased recording
$y_s\sim\text{Poisson}(\lambda_s g(a_s))$, $g$ $C^1$ increasing. A consistent
learner reaches the fixed point $\mu_s=\lambda_s g(\pi(\mu_s;M))$. Define
$\kappa_s=\big(\tfrac{\partial\log\pi}{\partial\log\mu}\big)\big(\tfrac{\partial\log g}{\partial\log a}\big)$.

\begin{theorem}[Amplification law]
\label{thm:law}
At any fixed point with $\kappa_s<1$,
$\dfrac{d\log\mu_s}{d\log\lambda_s}=\dfrac{1}{1-\kappa_s}$. In the homogeneous
case, $\mu_s/\mu_r=(\lambda_s/\lambda_r)^{1/(1-\kappa)}$, with a pole at
$\kappa=1$.
\end{theorem}
\noindent\emph{Proof.} Differentiate $\log\mu_s=\log\lambda_s+\log g(\pi(\mu_s))$
in $\log\lambda_s$; the second term contributes $\kappa_s\,d\log\mu_s/d\log\lambda_s$
by the chain rule; solve. Full proof and the coordinate-free (non-power-law)
verification in the supplement. $\square$

\begin{corollary}[Runaway-discrimination law]
Two groups with equal latent intensity but initial recording bias $b$ reach
recorded disparity $b^{1/(1-\kappa)}$ --- Ensign et al.'s runaway given an exact
form.
\end{corollary}

\begin{theorem}[Confidently wrong]
\label{thm:cw}
Let $\mathcal C_s$ be a conformal interval calibrated on observed residuals.
As $\kappa\uparrow$: (i) observed coverage $\mathbb P(y_s\in\mathcal C_s)\ge
1-\alpha$ is maintained; (ii) latent coverage $\mathbb P(\lambda_s\in\mathcal
C_s)$ collapses, since bias $|\mu_s-\lambda_s|$ grows while the half-width tracks
only Poisson dispersion; (iii) the biased world and an honest world
($\lambda'=\mu$, $\kappa=0$) are observationally identical, so no statistic of
recorded data distinguishes them.
\end{theorem}

% ============================================================
\section{Identification of the Feedback Gain}
\label{sec:ident}

By Theorem~\ref{thm:cw}(iii), $\kappa$ is not identified from passive data. But an
exogenous shock that scales the detection elasticity by $(1+\delta)$ in a treated
set $\mathcal T$ identifies it:

\begin{theorem}[Active identification]
With known policy elasticity $\beta$ and shock $\delta$, the difference-in-differences
$\mathrm{DiD}=\overline{\Delta\log\mu}_{\mathcal T}-\overline{\Delta\log\mu}_{\mathcal T^c}$
is a known strictly monotone function of $\kappa$; hence $\kappa=\Phi^{-1}(\mathrm{DiD})$,
without observing $\lambda$.
\end{theorem}

\noindent\textbf{Operating condition (discovered empirically, \S\ref{sec:exp}).}
The identifying shock must keep treated cells sub-runaway,
$\kappa(1+\delta)\le \kappa_{\max}<1$; otherwise the treated fixed point diverges
and identification degenerates. This is a genuine, citable design constraint.

% ============================================================
\section{Feedback-Corrected Conformal Prediction}
\label{sec:correct}

Given identified $\kappa$, the recording multiplier is $m_s=(\mu_s/M)^\kappa$ and
the latent rate is recovered by deflation $\hat\lambda_s=\mu_s/m_s$. We issue
count-distribution quantiles at $\hat\lambda_s$ (optionally widened by an
anchor-calibrated conformal margin), and \textbf{abstain} where $\kappa$ nears
the runaway threshold or the required deflation is extreme --- the honest failure
mode rather than the silent one.

\begin{algorithm}[h]
\caption{Feedback-corrected latent interval}
\begin{algorithmic}[1]
\REQUIRE recorded rates $\mu$, identified $\kappa$, level $\alpha$
\STATE $m_s \leftarrow (\mu_s/\text{mean}(\mu))^\kappa$;\quad $\hat\lambda_s \leftarrow \mu_s/m_s$
\STATE $[\ell_s,u_s]\leftarrow[\,Q_{\alpha/2}(\hat\lambda_s),\,Q_{1-\alpha/2}(\hat\lambda_s)\,]$
\STATE \textbf{if} $\kappa\ge\kappa_{\text{runaway}}$ \textbf{or} $m_s\notin[1/M_{\max},M_{\max}]$ \textbf{then} abstain on $s$
\end{algorithmic}
\end{algorithm}

% ============================================================
\section{Deployed Consequence: Feedback-Aware Safe Routing}
\label{sec:routing}

A crime-risk map is typically consumed downstream to \emph{route} people or
patrols. If routing plans against the recorded record, it will faithfully divert
civilians around historically over-recorded --- often over-policed ---
neighborhoods, laundering enforcement bias into navigation: the routing analogue
of the confidently-wrong phenomenon (``navigational redlining''). We make routing
the deployed consequence of the correction.

\textbf{Route on latent risk.} Node risk intervals are deflated to the latent
scale (\S\ref{sec:correct}) before edge costs are formed, so the router plans
against true risk, not enforcement artifacts.

\textbf{Tail-risk (CVaR) costs with abstention propagation.} Edge cost combines
distance with the $\beta$-CVaR of the corrected risk interval (the average of its
worst $(1-\beta)$ tail --- the robust choice for safety-critical routing); nodes
the correction flags un-correctable ($\kappa$ near runaway) incur an abstention
penalty, so the router avoids them where possible and the abstention monitor
triggers where not.

\textbf{Exposure-disparity audit.} We measure each demographic group's share of
routed risk-exposure against its share of latent incidence. Auditing the biased
vs.\ corrected risk field quantifies how much correction shrinks disparity. In
closed-loop simulation with a group structurally over-recorded by $1.8\times$,
the correction removes the \emph{feedback-induced} portion of the disparity:
worst-group exposure disparity falls from $0.29$ (biased) to $0.05$ at
$\kappa{=}0.85$ (Table~\ref{tab:routing}), more so as the loop strengthens. Prior
risk-aware routing plans on the observed field; routing on \emph{de-biased latent}
uncertainty with abstention propagation is, to our knowledge, new. (Implementation:
\texttt{src/civicsafe/routing/feedback\_aware.py}; experiment:
\texttt{scripts/routing\_disparity\_experiment.py}.)

\begin{table}[h]
\centering
\caption{Routing exposure disparity: biased vs.\ feedback-corrected risk field
(worst-group $|\text{disparity}|$, lower is fairer). Group 1 over-recorded
$1.8\times$.}
\label{tab:routing}
\begin{tabular}{cccc}
\toprule
$\kappa$ & biased & corrected & reduction \\
\midrule
0.00 & 0.287 & 0.287 & 0.000 \\
0.30 & 0.288 & 0.206 & 0.082 \\
0.50 & 0.294 & 0.148 & 0.147 \\
0.70 & 0.269 & 0.089 & 0.180 \\
0.85 & 0.237 & \textbf{0.045} & \textbf{0.193} \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================
\section{Deployment Safety: Live Monitor and Robustness Envelope}
\label{sec:deploy}

\textbf{Anytime-valid feedback tripwire.} Passive metrics cannot certify latent
coverage (Thm~\ref{thm:cw}), but a live monitor can detect \emph{drift}. We run a
test supermartingale over the coverage stream (method-of-mixtures betting;
Waudby-Smith \& Ramdas, 2024): under the null ``intervals are calibrated'' the
wealth is a non-negative supermartingale, so Ville's inequality gives a
\emph{time-uniform} false-alarm bound
$P(\exists t: W_t\!\ge\!1/\eta)\le\eta$ under continuous monitoring, with nothing
to tune. Empirically the tripwire holds its false-alarm budget on calibrated
streams, never fires under over-coverage, and fires early under a miscoverage
drift to $0.45$ (nominal $0.10$).

\textbf{Robustness envelope.} The correction uses an \emph{estimated} $\kappa$;
we report how much misspecification it tolerates. Latent coverage is
near-maximal at the true gain, and the \emph{robustness value} --- the largest
gain error keeping coverage above a floor --- shrinks as $\kappa\!\to\!1$: near
runaway the correction is more fragile, which sets the precision target for the
field identification of $\kappa$.

\textbf{Robustness to a misspecified recording model.} Beyond gain error, we
harden the correction against a wrong recording \emph{form}. Under a
Rosenbaum-style sensitivity model where the true recording multiplier lies within
a factor $\Gamma$ of the assumed one, a $\Gamma$-inflated corrected interval
covers the latent target for \emph{any} admissible recording model. Empirically
(true $\kappa{=}0.6$), un-inflated coverage degrades $0.95\!\to\!0.86$ as
$\Gamma$ grows to $2$, while the $\Gamma$-inflated interval stays $\ge 0.99$ at a
bounded $1.8\times$ width cost --- converting the method's key assumption into a
stated envelope. (Implementation: \texttt{feedback\_tripwire.py},
\texttt{sensitivity.py}, \texttt{correction\_robustness.py}; results in
\texttt{docs/RESULTS\_deployment\_layer.md},
\texttt{docs/RESULTS\_robustness\_and\_crosscity.md}.)

\textbf{Cross-city real data.} On the real Chicago (77 areas) and NYC (78
precincts) records (2018--2023), allocation on the raw recorded violent-crime
rate over-exposes the higher-minority stratum by $+39\%$ (Chicago) and $+31\%$
(NYC) relative to population share; correction at an assumed $\kappa{=}0.6$
attenuates this to $+16\%$ and $+12\%$. The pattern replicates across two
independent cities with real demographics. (Latent coverage is not validated on
real data --- the true rate is unobservable --- so this is a descriptive
companion to the simulation guarantee; script:
\texttt{scripts/cross\_city\_disparity.py}.)

% ============================================================
\section{Experiments}
\label{sec:exp}

\textbf{Closed-loop AOBF simulation.} Latent $\lambda\sim\text{Gamma}$, power-law
policy/detection, $\kappa$ identified by DiD (never observing $\lambda$), latent
coverage scored on fresh $\text{Poisson}(\lambda)$ draws (target $0.90$).

\begin{table}[h]
\centering
\caption{Feedback-corrected vs.\ naive latent coverage (target $0.90$). Naive
collapses as feedback rises; the correction holds throughout, abstaining on the
cells it cannot stand behind (kept fraction).}
\label{tab:correction}
\begin{tabular}{cccccc}
\toprule
$\kappa$ & shock $\delta$ & naive latent & $\hat\kappa$ (DiD) & \textbf{corrected latent} & kept frac. \\
\midrule
0.00 & 0.60 & 0.945 & 0.020 & \textbf{0.945} & 1.00 \\
0.30 & 0.60 & 0.901 & 0.300 & \textbf{0.957} & 1.00 \\
0.50 & 0.60 & 0.781 & 0.500 & \textbf{0.944} & 0.95 \\
0.70 & 0.29 & 0.526 & 0.700 & \textbf{0.926} & 0.57 \\
0.85 & 0.06 & 0.156 & 0.850 & \textbf{0.953} & 0.16 \\
\bottomrule
\end{tabular}
\end{table}

\noindent All theorems are verified numerically (\texttt{tests/test\_feedback\_law.py},
\texttt{tests/test\_latent\_correction.py}); the amplification law matches the
iterated dynamics to relative error $\le 10^{-4}$, including a non-power-law
(tanh policy, exponential detection) check.

\textbf{Limitations.} These are simulations of the AOBF model. External validity
requires the ShotSpotter/patrol-rollout DiD on real records, where $\kappa$ is
estimated from field data; this is the primary future-work item. The base ZINB/GNN
forecaster is applied prior art. The correction assumes the AOBF fixed-point form;
robustness to misspecified $g,\pi$ is bounded by the identification's operating
condition.

% ============================================================
\section{Conclusion}

Systems that choose where to look, then learn from what they see, can become
confidently wrong about the world they claim to measure --- and no passive metric
warns them. We gave this a closed form, an identification strategy, and, for the
first time, a correction that restores honest coverage of the latent process.
The next step is field identification of $\kappa$ from a natural experiment.

% ============================================================
\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{ensign2018} Ensign, D., Friedler, S.A., Neville, S., Scheidegger, C.,
Venkatasubramanian, S. (2018). Runaway feedback loops in predictive policing.
\textit{FAccT/PMLR 81}.

\bibitem{glaeser2003} Glaeser, E.L., Sacerdote, B.I., Scheinkman, J.A. (2003).
The social multiplier. \textit{Journal of the European Economic Association},
1(2--3), 345--353.

\bibitem{vanamsterdam2025} van Amsterdam, W.A.C., et al. (2025). When accurate
prediction models yield harmful self-fulfilling prophecies. \textit{Patterns}.

\bibitem{mendlerduenner2022} Mendler-D\"unner, C., et al. (2022). Anticipating
performativity by predicting from predictions. \textit{NeurIPS}.

\bibitem{gibbs2021} Gibbs, I., Cand\`es, E.J. (2021). Adaptive conformal
inference under distribution shift. \textit{NeurIPS}.

\bibitem{romano2019} Romano, Y., Patterson, E., Cand\`es, E.J. (2019).
Conformalized quantile regression. \textit{NeurIPS}.

\bibitem{zhuang2022} Zhuang, D., et al. (2022). Uncertainty quantification of
sparse travel demand prediction with spatial-temporal graph neural networks
(STZINB-GNN). \textit{KDD}.

\bibitem{wang2024} Wang, et al. (2024). Uncertainty-aware crime prediction with
spatial-temporal multivariate ZINB graph neural networks (STMGNN-ZINB).

\bibitem{perdomo2020} Perdomo, J.C., Zrnic, T., Mendler-D\"unner, C., Hardt, M.
(2020). Performative prediction. \textit{ICML}.

\bibitem{hersbach2000} Hersbach, H. (2000). Decomposition of the CRPS for
ensemble prediction systems. \textit{Weather and Forecasting}, 15(5).

\end{thebibliography}

\end{document}
